Naturalness in U(1) extensions of the MSSM

### 2014-12-18 paper/e6ssm-tuning-paper.tex
fixed merge problem that I missed.  ugh..
Moved definition of d coefficient to EWSB section, removed additional discussion from results section

Written in the form of Eq. (\ref{eq:E6MZequation}) we see the potential new source of fine tuning alluded to above, in the form of the third term on the right-hand side. For large values of the VEV $s$, the $D$-term contributions can be quite a bit larger than $M_Z^2$. In particular, recent experimental limits \cite{Aad:2014cka} require that the $Z'$ mass be large, with bounds of $M_{Z'}\gtrsim 2.51$ TeV in $U(1)_\psi$ models and $M_{Z'}\gtrsim 2.62$ TeV in $U(1)_\chi$ models. To satisfy these limits typically requires large values of the singlet VEV $s$. For example $s\gtrsim 6$ TeV is required in the E$_6$SSM with $U(1)'=U(1)_N$, so that $|D_{H_u}|,|D_{H_d}|\gg M_Z^2$ for $E_6$ models with $\tilde{Q}_S\neq 0$. As a result the remaining terms on the right-hand side of Eq. (\ref{eq:E6MZequation}) must be tuned to cancel the resulting very large contribution to $M_Z$. Moreover, because this is a large \textit{tree level} fine tuning, it may negate the increase in naturalness that is associated with having a reduced need for heavy superpartners.

@@

To illustrate this, in Fig. \ref{} we show the variation in fine
tuning when we fix $\mu=-97$ GeV and $B=-85$ GeV, and scan over the
stop masses and mixing, with $500 \textrm{ GeV} \leq
m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and $-3810\textrm{ GeV}\leq
A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV, $50$ TeV and $100$ TeV.
However this tuning only really becomes large for very large stop
masses.  On the other hand if cuttoff is increased then stop masses of
several TeV can result in tuning of [insert] as shown in
Fig.~\ref{}.\\ \\

@@

The exact level of tuning from the $Z^\prime$ depends on the charges of
the extra $U(1)$ gauge symmetry it is associated with.  In Fig \ref{}
we look at the fine tuning for other U(1) extensions for the same
$Z^\prime$ masses as we did for the E$_6$SSM.
 % - if room, plots for different U(1)' charge Here we see that the As
can be seen in Fig.~\ref{} the severity of the tunings varies quite a
bit.  This is because the charges appear as coefficients in front of
the $Z^\prime$ mass in the EWSB condition.  These charges change the
value of the coefficient $d$ in Eq.~\ref{}.  The values of the
coefficient $d$ in each model is $\{-0.01, 0.40, 0.50, 0.81 \}$ for
$\{U(1)_I, U(1)_N, U(1)_\psi, U(1)_\eta \}$ and this determines which
of the models are most tuned.

>>>>

Written in the form of Eq. (\ref{eq:E6MZequation}) we see the potential new source of fine tuning alluded to above, in the form of the third term on the right-hand side. For large values of the VEV $s$, the $D$-term contributions can be quite a bit larger than $M_Z^2$. In particular, recent experimental limits \cite{Aad:2014cka} require that the $Z'$ mass be large, with bounds of $M_{Z'}\gtrsim 2.51$ TeV in $U(1)_\psi$ models and $M_{Z'}\gtrsim 2.62$ TeV in $U(1)_\chi$ models. To satisfy these limits typically requires large values of the singlet VEV $s$. For example $s\gtrsim 6$ TeV is required in the E$_6$SSM with $U(1)'=U(1)_N$, so that $|D_{H_u}|,|D_{H_d}|\gg M_Z^2$ for $E_6$ models with $\tilde{Q}_S\neq 0$. As a result the remaining terms on the right-hand side of Eq. (\ref{eq:E6MZequation}) must be tuned to cancel the resulting very large contribution to $M_Z$. Moreover, because this is a large \textit{tree level} fine tuning, it may negate the increase in naturalness that is associated with having a reduced need for heavy superpartners. In $U(1)$ extended models for which $\tilde{Q}_S\neq 0$, the importance of the $Z^\prime$ mass to the fine tuning in these models can be made even clearer by writing Eq. (\ref{eq:E6MZequation}) in the form given in \cite{Athron:2013ipa},
\begin{equation}\label{eq:originalE6MZequation}
c(\theta,\tan\beta)\frac{M_Z^2}{2}=-\frac{\lambda^2 s^2}{2}+\frac{\tilde{m}_{H_d}^2-\tilde{m}_{H_u}^2\tan^2\beta}{\tan^2\beta-1}+d(\theta,\tan\beta)\frac{M_{Z^\prime}^2}{2}
c(\theta,\tan\beta)&=1-\frac{4}{\left ( \tan^2\beta-1\right )}\frac{g_1'^2}{\bar{g}^2}\left ( \tilde{Q}_1-\tilde{Q}_2\tan^2\beta \right )\left ( \tilde{Q}_1\cos^2\beta+\tilde{Q}_2\sin^2\beta \right ),\label{eq:cdefn}\\
d(\theta,\tan\beta)&= \frac{\tilde{Q}_1-\tilde{Q}_2\tan^2\beta}{\tilde{Q}_S\left ( \tan^2\beta - 1\right )}.\label{eq:ddefn}
Written like so it is evident that the fine tuning contribution coming from the new $D$-terms depends both on the $U(1)'$ charges and the $Z^\prime$ mass, and that the tuning can be expected to increase with $M_{Z^\prime}$. As shall be shown below, the exact size of this tuning then depends strongly on the choice of $U(1)'$ charges, via the coefficient $d$. 

@@

To illustrate this, in Fig. \ref{} we show the variation in fine
tuning when we fix $\mu=-97$ GeV and $B=-85$ GeV, and scan over the
stop masses and mixing, with $500 \textrm{ GeV} \leq
m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and $-3810\textrm{ GeV}\leq
A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV, $50$ TeV and $100$ TeV.
However this tuning only really becomes large for very large stop
masses.  On the other hand if the cuttoff is increased then stop masses of
several TeV can result in tuning of [insert] as shown in
Fig.~\ref{}.\\ \\

@@

The exact level of tuning from the $Z^\prime$ depends on the charges of
the extra $U(1)$ gauge symmetry it is associated with.  In Fig \ref{}
we look at the fine tuning for other U(1) extensions for the same
$Z^\prime$ masses as we did for the E$_6$SSM.
% - plots for different U(1)' charge

% DH:: for reference:
%      U(1)_I \theta = \pi+\arctan\sqrt{3/5}
%      U(1)_\psi \theta = \pi / 2
%      U(1)_\eta \theta = -\arctan\sqrt{5/3}

As can be seen in Fig.~\ref{} the severity of the tunings varies quite a
bit.  This is because the charges appear as coefficients in front of
the $Z^\prime$ mass in the EWSB condition.  These charges change the
value of the coefficient $d$ in Eq. (\ref{eq:originalE6MZequation}).  The values of the
coefficient $d$ in each model, for $\tan\beta = 10$, is $\{-0.01, 0.40, 0.50, 0.81 \}$ for
$\{U(1)_I, U(1)_N, U(1)_\psi, U(1)_\eta \}$ and this determines which
of the models are most tuned.
